---
title: "Subinhibitory Concentration of Antibiotics (SICA)"
author: "Molly Kuhs"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: True
    toc_depth: 4
    number_sections: False
    highlight: pygments
    theme: cosmo
---

<style type="text/css">

body{ /* Normal  */
   font-size: 14px;
}
td {  /* Table  */
   font-size: 12px;
}
h1 { /* Header 1 */
 font-size: 26px;
 font-weight: bold;
}
h2 { /* Header 2 */
 font-size: 26px;
 font-weight: bold;
}
h3 { /* Header 3 */
 font-size: 18px;
 font-weight: bold;
}
h4 { /* Header 4 */
 font-size: 16px;
 font-weight: bold;
}
code.r{ /* Code block */
  font-size: 12px;
}
pre { /* Code block */
  font-size: 12px
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #848587;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

Data Availability: 

Full dataset, code, and figures can be found here: https://github.com/makuhs/SICA

Note: Included code creates base versions of all published figures. Additional elements such as legends, titles, and annotations were added via post-processing in Adobe Illustrator. 

## Packages
```{r}
library(tidyverse)
library(multcomp)
library(vegan)
library(ggpubr)
library(dplyr)
library(vcd)
```

## Data 
### Data Load
```{r}
data <- read.csv("fullBiologSICA.csv")%>%
  subset(select = -X )
```

### Nutrient Use Profiles
```{r}
metrics <- data %>%
  mutate(niche.width = rowSums(.[5:99]!=0))%>%
  mutate(total.growth = rowSums(.[5:99]))%>%
  mutate(growth.eff = .$total.growth / .$niche.width)%>%
  dplyr::select(Plot, Letter,SICA, Isolate, niche.width, total.growth, growth.eff)

metricControl <- metrics %>%
  filter(SICA == "Control")%>%
  subset(select = -SICA)%>%
  rename_with(~c("nw_c", "tg_c", "ge_c"), c(niche.width, total.growth, growth.eff))

metricSICA <- metrics %>%
  filter(SICA != "Control")

metrics2 <- merge(metricSICA, metricControl, by=c("Plot", "Letter", "Isolate")) %>%
  group_by(SICA) %>%
  filter (niche.width > 25 & total.growth >= 2) # remove isolates with extreme low growth

#calculate difference from control 
diff <- metrics2 %>%
  mutate(nw_diff = (niche.width - nw_c),
         tg_diff = (total.growth - tg_c),
         ge_diff = (growth.eff - ge_c)) 

diffSum <- metrics %>%
  filter(niche.width >= 25 & total.growth >2)

```

## Main 

### Figure 1
```{r}
#Nutrient Use Phenotypes
diffSum2 <- diffSum %>%
  mutate(Treatment = case_when(Plot == "3" ~ "LNI",
                               Plot == "4" ~ "HNI"))

diffSum2$SICA <- factor(diffSum2$SICA, levels = c("Control", "Streptomycin", "Vancomycin", "Streptothricin", "Rifampicin", "Tetracycline", "Chloramphenicol"))

diffSum2$Treatment <- factor(diffSum2$Treatment, levels = c("LNI", "HNI"))

# categorize difference in presence of SICA as increasing, decreasing or no change from control 
subset <- diff %>%
  mutate(nw_direction = case_when(nw_diff > 0 ~ "I", 
                                  nw_diff < 0 ~ "D", 
                                  nw_diff == 0 ~"NC"),
         tg_direction = case_when(tg_diff > 0 ~ "I", 
                                  tg_diff < 0 ~ "D", 
                                  tg_diff == 0 ~"NC"),
         ge_direction = case_when(ge_diff > 0 ~ "I", 
                                  ge_diff < 0 ~ "D", 
                                  ge_diff == 0 ~"NC"),
         Treatment = case_when(Plot == "3" ~ "LNI",
                               Plot == "4" ~ "HNI")) 

subset <- subset[,-8:-12]
subset$Treatment <- factor(subset$Treatment, levels = c("LNI", "HNI"))

controlSub <- diffSum2 %>%
  filter(SICA == "Control")


color <- c("grey80", 
           "#c7dbed",
           "#c7dbed", 
           "#c7dbed", 
           "#c7dbed", 
           "#c7dbed", 
           "#c7dbed",
           "#36688a", #for points: decrease with SICA
           "#a6453d", #for points: increase with SICA
           "grey65") #for porints: sig. change with SICA

fill <- c("grey70", "#5499C7","#5499C7","#5499C7","#5499C7","#5499C7","#5499C7")
```

#### 1A
```{r}

#Niche Width
a1<- ggplot(diffSum2, aes(SICA, niche.width, fill = SICA, color = SICA))+
  geom_boxplot(outlier.shape = NA, alpha=0.2, size=0.6)+ 
  geom_point(aes(SICA, niche.width), size = 1, alpha=0.6, data=controlSub)+
  geom_point(aes(SICA, niche.width, color = nw_direction), size = 1, alpha=0.6, data = subset)+
  labs(x = "",
       y= "Niche Width\n")+
  theme_minimal()+
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        panel.spacing = unit(2.5, "lines"),
        axis.title=element_text(size=9),
        strip.text.x=element_text(face = "bold"))+
  facet_wrap(~Treatment)+
  scale_color_manual(values=color)+
  scale_fill_manual(values=fill)

#LNI niche width ANOVA 
modnw3 <- diffSum2 %>%
  filter(Plot == 3) %>%
  aov(niche.width ~ SICA, data=.)

summary(modnw3)

#HNI niche width ANOVA 
modnw4 <- diffSum2 %>%
  filter(Plot == 4) %>%
  aov(niche.width ~ SICA, data=.)

summary(modnw4)
TukeyHSD(modnw4)

modnw4Comp <- glht(modnw4, linfct=mcp(SICA="Tukey"))
cld(modnw4Comp)

```

#### 1B
```{r}

#Total Growth 
b1<- ggplot(diffSum2, aes(SICA, total.growth, fill = SICA, color = SICA))+
  geom_boxplot(outlier.shape = NA, alpha=0.2, size=0.6)+ 
  geom_point(aes(SICA, total.growth), size = 1, alpha=0.6, data=controlSub)+
  geom_point(aes(SICA, total.growth, color = tg_direction), size = 1, alpha=0.6, data = subset)+
  labs(x = "",
       y= "Total Growth\n")+
  theme_minimal()+
  facet_wrap(~Treatment)+
  theme(legend.position = "none",
        panel.spacing = unit(2.5, "lines"),
        axis.title=element_text(size=9),
        strip.text.x=element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))+
  scale_color_manual(values=color)+
  scale_fill_manual(values=fill)

#LNI total growth ANOVA 
modtg3 <- diffSum2 %>%
  filter(Plot == 3) %>%
  aov(total.growth ~ SICA, data=.)
summary(modtg3)

#HNI total growth ANOVA 
modtg4 <- diffSum2 %>%
  filter(Plot == 4) %>%
  aov(total.growth ~ SICA, data=.)

summary(modtg4)
TukeyHSD(modtg4)

modtg4Comp <- glht(modtg4, linfct=mcp(SICA="Tukey"))
cld(modtg4Comp)

```

#### 1C
```{r}
biologMap <- data
set.seed(10194)

#LNI  
bio3 <- biologMap %>%
  filter(Plot == "3")

data3 <- bio3[5:ncol(bio3)]
p3Matrix <- as.matrix(data3)
p3Meta <- bio3[1:4]

bio.3.dist <- vegdist(p3Matrix, method='bray')

adonis2(bio.3.dist~p3Meta$SICA)

bd3<- betadisper(bio.3.dist, p3Meta$SICA)
anova(bd3)
permutest(bd3)
```

```{r, results = "hide"}
bio.3.nmds <- metaMDS(bio.3.dist,
                      distance = 'bray',
                      k=2, try=100, trymax=250)
```

```{r}
bio.3.nmds.points <- data.frame(bio.3.nmds$points)
bio.3.nmds.points$SICA <- as.factor(bio3$SICA)

bio.3.nmds.points$SICA <- factor(bio.3.nmds.points$SICA, levels = c("Control", "Chloramphenicol", "Rifampicin", "Streptomycin", "Streptothricin", "Tetracycline", "Vancomycin"))

bio3C <- bio.3.nmds.points %>%
  filter(SICA == "Control")

nds1<- bio.3.nmds.points %>%
  filter(SICA != "Control") %>%
  ggplot(., aes(x=MDS1, y=MDS2, color = SICA)) +
  geom_point(data = bio3C, aes(MDS1, MDS2), size = 1)+
  stat_ellipse(data = bio3C, aes(MDS1, MDS2), geom = "polygon", fill = "grey65", alpha=0.2, linetype = 2, show.legend = F)+
  stat_ellipse(show.legend = F)+
  geom_point(size = 1, alpha=0.8)+
  scale_color_manual(values = c("grey65", #Control
                                "#ee8e7d", #Chlor
                                "#f7c981", #Rif
                                "#add193", #Streptom
                                "#c2e8e8", #Streptot
                                "#76add2", #Tet
                                "#c0b8d4"))+ #Vanc
  theme_minimal() +
  theme(legend.position = "none",
        axis.title=element_text(size=9))+
  labs(x='\nMDS1', 
       Y='MDS2\n', 
       color='')+
  ylim(-1.3, 1.3)+
  xlim(-1.8, 1.8)

#HNI 
bio4 <- biologMap %>%
  filter(Plot == "4")

data4 <- bio4[5:ncol(bio4)]
p4Matrix <- as.matrix(data4)
p4Meta <- bio4[1:4]

bio.4.dist <- p4Matrix %>%
  vegdist(method='bray')

adonis2(bio.4.dist~p4Meta$SICA)

bd4 <- betadisper(bio.4.dist, p4Meta$SICA)
permutest(bd4)
```

```{r, results = "hide"}
bio.4.nmds <- metaMDS(bio.4.dist,
                      distance = 'bray',
                      k=2, try=100, trymax=250)
```

```{r}
bio.4.nmds.points <- data.frame(bio.4.nmds$points)
bio.4.nmds.points$SICA <- as.factor(bio4$SICA)

bio.4.nmds.points$SICA <- factor(bio.4.nmds.points$SICA, levels = c("Control", "Chloramphenicol", "Rifampicin", "Streptomycin", "Streptothricin", "Tetracycline", "Vancomycin"))

bio4C <- bio.4.nmds.points %>%
  filter(SICA == "Control")

nds2<- bio.4.nmds.points %>%
  filter(SICA != "Control") %>%
  ggplot(., aes(x=MDS1, y=MDS2, color = SICA)) +
  geom_point(data = bio4C, aes(MDS1, MDS2), size = 1, show.legend = F)+
  stat_ellipse(data = bio4C, aes(MDS1, MDS2), geom = "polygon", fill = "grey65", alpha=0.2, linetype = 2, show.legend = F)+
  stat_ellipse(show.legend = F)+
  geom_point(size = 1, alpha=0.7)+
  scale_color_manual(values = c("grey65",
                                "#ee8e7d", 
                                "#f7c981", 
                                "#add193",
                                "#c2e8e8",
                                "#76add2",
                                "#c0b8d4"))+
  theme_minimal() +
  labs(x='\nMDS1', 
       Y='MDS2\n', 
       color='')+
  theme(legend.position = "none",
        axis.title=element_text(size=9))+
  ylim(-1.3, 1.3)+
  xlim(-1.8, 1.8)

# combine panels 
c1<- ggarrange(nds1, NULL, nds2,
               nrow = 1, 
               widths = c(1, 0.05, 1))
```

#### Final
```{r, fig.dim = c(5, 7)}
finalfig1 <- ggarrange(a1, b1, c1,
                       nrow = 3,
                       heights = c(1.2, 1.5, 1.2))
finalfig1

ggsave("finalfig1.pdf", width = 5.5, height = 7.7, units = "in")

```


### Figure 2 

```{r}
biolog_long <- data %>%
  pivot_longer(cols = 'B1':'H12',
               names_to = "BiologWell", 
               values_to = "OD", 
               values_drop_na = TRUE)

#add specific carbon source information to growth data
c_class <- read.csv("Biolog_Nutrient_List.csv")
cClassTable <- c_class[,c(1,2)]

nutCat <- merge(biolog_long, c_class, by="BiologWell")%>%
  mutate(nutCount = case_when(OD>0~"1",
                              OD<!0~"0"))
nutCat$nutCount <- as.numeric(nutCat$nutCount)
nutCat$Plot <- as.factor(nutCat$Plot)

#calculate difference in growth between control and SICAs
control <- nutCat %>%
  filter(SICA == "Control")

nutChange <- merge(nutCat, control, by=c("Isolate", "Nutrient", "BiologWell", "Plot", "Letter", "CarbonSource", "CarbonType", "Count", "NumType"))%>%
  rename(SICA = SICA.x,
         Control = SICA.y,
         ODs = OD.x,
         ODc = OD.y) %>%
  mutate(Change = ODs - ODc,
         absChange = abs(Change),
         relChange = round(Change/ODc,2))


#determine 'significant change' threshold 
nutChange %>%
  summarise(mean = mean(absChange),
            sd=sd(absChange),
            n=n(),
            se=sd/sqrt(n))

#define change direction and significance by standard deviation (change > +/- 0.06OD)
nutChangeCat <- nutChange %>%
  mutate(nutCat = case_when(ODs==0 & ODc > 0 ~"supp",
                            ODs > 0 & ODc == 0 ~ "induc",
                            ODs > 0 & ODc > 0 & Change >= 0.06 ~ "incre",
                            ODs > 0 & ODc > 0 & Change <= -0.06 ~ "decre",
                            ODs >= 0 & ODc >= 0 & absChange < 0.06 ~ "no_grow"))%>%
  filter(SICA != "Control")

nutChangeCat$nutCat <- factor(nutChangeCat$nutCat, levels = c("supp", "decre", "incre", "induc", "no_grow"))
nutChangeCat$SICA <- factor(nutChangeCat$SICA, levels = c("Chloramphenicol", "Tetracycline", "Rifampicin", "Streptomycin", "Vancomycin", "Streptothricin"))

#responsiveness t-test: 
t.test(absChange ~ Plot, data = nutChangeCat)


```

```{r,fig.dim=c(3.55, 5.5)}
#LNI

nutODLim3 <- nutCat %>%
  filter(Plot == "3",
         SICA == "Control") %>%
  group_by(CarbonSource) %>%
  summarise(mean = mean(OD))

nutChangeCat3 <- nutChangeCat %>%
  filter(Plot == "3")

#manually remove isolates with niche.width > 25 & total.growth >= 2
nutODLim3combo <- merge(nutChangeCat3, nutODLim3, by = "CarbonSource")%>%
  filter(Isolate != "3B" | SICA != "Chloramphenicol") %>%
  filter(Isolate != "3C" | SICA != "Chloramphenicol") %>%
  filter(Isolate != "3C" | SICA != "Tetracycline") %>%
  filter(Isolate != "3E" | SICA != "Streptothricin") %>%
  filter(Isolate != "3G" | SICA != "Chloramphenicol") %>%
  filter(Isolate != "3F" | SICA != "Streptothricin")

#summarise across all isolates 
mapSum3 <- nutODLim3combo %>%
  group_by(SICA, nutCat)%>%
  summarise(count = n())%>%
  ungroup()%>%
  group_by(SICA) %>%
  mutate(total = sum(count),
         prop = count/total)

mapSum3$prop <- round(mapSum3$prop, 2)
mapSum3$nutCat <- factor(mapSum3$nutCat, levels = c("supp", "decre", "incre", "induc", "no_grow"))

F2A<- mapSum3 %>%
  filter(nutCat != "no_grow")%>%
  ggplot(., aes(nutCat, prop, fill = nutCat))+
  geom_col(alpha=0.85)+
  facet_wrap(~SICA, nrow=6)+
  scale_fill_manual(values = c("#EA725D", "#F5BB62", "#85D1D1", "#5499C7"),
                    labels = c("Suppressed Growth", "Reduced Growth", "Increased Growth", "Induced Growth"),
                    name = "")+
  theme_minimal()+
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 8),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 8))+
  labs(x = "",
       y = "Freq. of change in isolate growth with SICA\n")+
  scale_y_continuous(limits = c(0, 0.45), breaks = seq(0, 0.45, by = 0.2))

```

```{r, fig.dim=c(3.55, 5.5)}
# HNI 

nutODLim4 <- nutCat %>%
  filter(Plot == "4",
         SICA == "Control") %>%
  group_by(CarbonSource) %>%
  summarise(mean = mean(OD))

nutChangeCat4 <- nutChangeCat %>%
  filter(Plot == "4")

#manually remove isolates with niche.width > 25 & total.growth >= 2
nutODLim4combo <- merge(nutChangeCat4, nutODLim4, by = "CarbonSource")%>%
  filter(Isolate != "4E" | SICA != "Chloramphenicol") 

#summarise across all isolates 
mapSum4 <- nutODLim4combo %>%
  group_by(SICA, nutCat)%>%
  summarise(count = n())%>%
  ungroup()%>%
  group_by(SICA) %>%
  mutate(total = sum(count),
         prop = count/total)

mapSum4$prop <- round(mapSum4$prop, 2)
mapSum4$nutCat <- factor(mapSum4$nutCat, levels = c("supp", "decre", "incre", "induc", "no_grow"))

F2B<- mapSum4 %>%
  filter(nutCat != "no_grow")%>%
  ggplot(., aes(nutCat, prop, fill = nutCat))+
  geom_col(alpha=0.85)+
  facet_wrap(~SICA, nrow=6)+
  scale_fill_manual(values = c("#EA725D", "#F5BB62", "#85D1D1", "#5499C7"),
                    labels = c("Suppressed Growth", "Reduced Growth", "Increased Growth", "Induced Growth"),
                    name = "")+
  theme_minimal()+
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 8),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 8))+
  labs(x = "",
       y = "Freq. of change in isolate growth with SICA\n")+
  scale_y_continuous(limits = c(0, 0.45), breaks = seq(0, 0.45, by = 0.2))

ggarrange(F2A, NULL, F2B,
          nrow = 1, 
          widths = c(1, 0.1, 1))

ggsave("Fig2.pdf", width = 3.55, height = 5.5, units = "in", device = cairo_pdf)

```


### Figure 3 
```{r, fig.dim=c(6, 4)}

# Escape Ratio 
escape <- read.csv("escape_ratio.csv")

escapeLong <- escape %>%
  pivot_longer(cols = c(alt_escape_ratio_x_from_y, alt_escape_ratio_y_from_x), names_to = "direction", values_to = "escapeRatio")%>%
  mutate(target = case_when(direction == "alt_escape_ratio_x_from_y" ~ Isolate_x, direction == "alt_escape_ratio_y_from_x" ~Isolate_y),
         pairing = case_when(direction == "alt_escape_ratio_x_from_y" ~ Isolate_y, direction == "alt_escape_ratio_y_from_x" ~Isolate_x))

cleanLong <- escapeLong[,-2:-3]

cleanLong <- cleanLong[,-3:-4] %>%
  mutate(log = log(escapeRatio),
         logE = case_when(log == "-Inf" ~ 0,
                          log != "-Inf" ~ log))


##categorize change in growth as increase or decreasing using 10% cut off for "change" 

split <- cleanLong %>%
  mutate(comp = case_when(logE >= 0.1 ~ "decrease", 
                          #decrease comp (inc. CFG)
                          logE <= -0.1 ~ "increase", 
                          # increase comp (dec CFG)
                          logE > -0.1 & logE < 0.1 ~ "no change"))
```

#### 3B
```{r}
split %>%
  filter(comp != 'no change')%>%
ggplot(., aes(x=logE, fill = comp))+
  geom_histogram(alpha=0.8, color =" white", size=0.2, breaks = seq(-6, 6, 0.5), aes(y=..count../sum(..count..)))+
  geom_vline(xintercept = 0, color = "white", size = 0.75)+
  geom_vline(xintercept = 0, lty = 3, color = "grey30")+
  theme_minimal()+
  facet_wrap(~Plot)+
  labs(x = "",
       y = "proportion of interactions\n",
       fill = "")+
  scale_fill_manual(values = c("#76add2", "#ee8e7d"),
                    labels = c("decrease with SICA", "increase with SICA")
                    )+
  theme(legend.position = "none")+
  scale_y_continuous(breaks = seq(0, 0.1, by = 0.02))

ggsave("Fig3.pdf", width = 6, height = 4, units = "in", device = cairo_pdf)


#chi-squared test difference between Plot: 
splitSum <- split %>%
  group_by(Plot, comp)%>%
  summarise(count = n())%>%
  arrange(comp)

chiPlot <- matrix(splitSum$count,
               nrow = 3, 
               byrow = TRUE, 
               dimnames = list(c("decreasedComp", "increaseComp", "noChange"), 
                               c("LNI", "HNI")))
chiPlot

chisq.test(chiPlot)

chiPlotPH <- chisq.test(chiPlot)
residuals(chiPlotPH, type = "standardized")

```

#### 3C
```{r}

splitSumDiff<- split %>%
  group_by(Plot, SICA, comp)%>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(Plot, SICA) %>%
  mutate(total = sum(count),
         change = round((count/total)*100,1))

#chi squared plot difference chloramphenicol 
splitSumChlor <- split %>%
  filter(SICA == "Chloramphenicol",
         comp != "no change")%>%
  group_by(Plot, comp)%>%
  summarise(count = n())%>%
  arrange(comp)

chiChlor <- matrix(splitSumChlor$count, 
               nrow = 2, 
               byrow = TRUE, 
               dimnames = list(c("decreaseComp", "increaseComp"), 
                               c("LNI", "HNI")))
chiChlor 

chisq.test(chiChlor)

#chi squared plot difference tetracycline 
splitSumTet <- split %>%
  filter(SICA == "Tetracycline",
         comp != "no change")%>%
  group_by(Plot, comp)%>%
  summarise(count = n()) %>%
  arrange(comp)

chiTet <- matrix(splitSumTet$count, 
               nrow = 2, 
               byrow = TRUE, 
               dimnames = list(c("decreaseComp", "increaseComp"), 
                               c("LNI", "HNI")))
chiTet
chisq.test(chiTet)

```


### Figure 4
```{r, fig.dim=c(8,4)}
altEscape <- escape [,-5:-6] 

logEscape <- altEscape %>%
  mutate(logX = log(alt_escape_ratio_x_from_y),
         logY = log(alt_escape_ratio_y_from_x)) %>%
  filter(logX != "-Inf")%>%
  filter(logY != "-Inf")%>%
  mutate(direct=case_when(
    logX >= 0.09 & logY >= 0.09 ~ "++",
    logX <= -0.11 & logY <= -0.11 ~ "--",
    logX >= 0.09 & logY <= -0.11 ~ "+-",
    logY >= 0.09 & logX <= -0.11 ~ "+-",
    logY < 0.09 & logY > -0.11 & logX < 0.09 & logX > -0.11~"00",
    logY < 0.09 & logY > -0.11 & logX >= 0.09 ~ "+0",
    logX < 0.09 & logX > -0.11 & logY >= 0.09 ~ "+0",
    logY < 0.09 & logY > -0.11 & logX <= -0.11 ~ "-0",
    logX < 0.09 & logX > -0.11 & logY <= -0.11 ~ "-0"
  ))
#log 1.1 or 0.9

escapeSum <- logEscape %>%
  group_by(Plot, SICA, direct)%>%
  summarise(counts = n())

escapeSum2 <- escapeSum %>%
  group_by(Plot, SICA)%>%
  mutate(total = sum(counts)) %>%
  ungroup() %>%
  mutate(percent = (counts/total)*100) %>%
  mutate(Treatment = case_when(Plot == "3" ~ "LNI",
                               Plot == "4" ~ "HNI"))

escapeSum2$direct <- factor(escapeSum2$direct, levels = c("++", "+0", "+-", "-0", "--","00"))

escapeSum2$Treatment <- factor(escapeSum2$Treatment, levels = c("LNI", "HNI"))

escapeSum2$SICA <- factor(escapeSum2$SICA, levels = c("Chloramphenicol", "Tetracycline", "Streptomycin", "Streptothricin", "Rifampicin", "Vancomycin"))
  
escapeSum2 %>%
  ggplot(., aes(direct, percent, fill = direct))+
  geom_col(alpha=0.9)+
  facet_grid(Treatment~SICA)+
  scale_fill_manual(values = c("#76add2", "#CEE9E9", "#A0CECE", "#F6D283", "#E79789", "grey85"))+
  scale_x_discrete()+
  theme_minimal()+
  theme(legend.position = "none",
        panel.spacing = unit(0.5, "lines"),
        axis.text.x=element_text(size=6.5),
        axis.title.y=element_text(size=10),
        strip.text.x=element_text(face = "bold"))+
  labs(x = "",
       y = "Frequency (%)\n")

ggsave("Fig4.pdf", width = 8, height = 4, units = "in", device = cairo_pdf)

escapeStats <- logEscape %>%
  group_by(Plot, direct)%>%
  summarise(counts = n())

new_row <- data.frame(Plot = 4, direct = "++", counts = 0) 
#add row for plot 4 ++ as 0

escapeStatsFull <- rbind(escapeStats, new_row)%>%
  arrange(direct)

escapeChi <- matrix(escapeStatsFull$counts,
               nrow = 6, 
               byrow = TRUE, 
               dimnames = list(c("++", "+-", "+0", "--", "-0", "00"), 
                               c("LNI", "HNI")))

escapeChi

chiq<- chisq.test(escapeChi, simulate.p.value = TRUE) 
#simulate due to low counts
chiq

residuals(chiq, type = "standardized")

```



## Supplement 

#### Table S1
```{r}
#soil data
soilData <- read.csv("SoilData.csv")%>%
  filter(LabPlot_ID == 3| LabPlot_ID == 4)

soilSum <- soilData %>%
  group_by(LabPlot_ID)%>%
  summarise(pH = mean(pH),
            BrayP = mean(Bray.P),
            K = mean(K),
            totalC = mean(Total.C),
            NO3N = mean(NO3.N))

#t-test: 
t.test(pH ~ LabPlot_ID, data=soilData)

```

#### Table S2
```{r}
#MIC values
rawData <- read.csv("MICSICA_data.csv")%>%
  filter(Value == "MIC")%>%
  pivot_longer(cols = Streptomycin:Streptothricin,
               names_to = "antibio",
               values_to = "MIC")%>%
  mutate(plot = str_sub(Isolate, 1, 1))

#MIC value anova
mod3<- aov(MIC~plot, data = rawData)
summary(mod3)
```

#### Figure S1
```{r}

#niche width difference w/ and w/o SICA
diffSum %>%
  mutate(group = case_when(SICA == "Control" ~ "Control",
                           SICA != "Control" ~ "With SICA"))%>%
  ggplot(., aes(group, niche.width, fill = as.factor(Plot)))+
  geom_bar(stat = "summary",
           width = 0.7,
           position = position_dodge(width = 0.8), 
           alpha=0.8)+
  geom_errorbar(stat = "summary",
                position = position_dodge(width = 0.8),
                width = 0.1,
                show.legend = F)+
  scale_fill_manual(values = c("#ba565c", "#65a4c2"),
                    labels = c("Low Nutrient Isolates", 
                               "High Nutrient Isolates"),
                    name = "")+
  theme_minimal()+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")+
  labs(x = "",
       y = "Niche Width\n")+
  coord_cartesian(ylim = c(65, 90))

# niche width t.test without SICA
diffSum %>%
  filter(SICA == "Control")%>%
  t.test(niche.width ~ Plot, data = .)

# niche width t.test with SICA
diffSum %>%
  filter(SICA != "Control")%>%
  t.test(niche.width ~ Plot, data = .)

```

#### Figure S2
```{r, fig.dim = c(8.5, 9)}
#LNI
#y-axis ordered by mean growth in control 
ggplot(nutODLim3combo, aes(Letter, reorder(CarbonSource, mean), fill=nutCat))+
  geom_tile()+
  facet_wrap(~SICA, nrow = 1, scales = "free_x")+
  scale_fill_manual(values = c("#EA725D","#F5BB62", "#85D1D1", "#5499C7", "grey99"),
                    labels = c("Suppressed Growth","Reduced Growth", "Increased Growth", "Induced Growth", "No Change"),
                    name = "")+
  labs(x="",
       y="")+
  theme_light()+
  theme(legend.position = "none",
        legend.text = element_text (size=8),
        axis.title.x = element_text (size = 8), 
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(), 
        strip.text.y.left = element_text(angle = 0),
        strip.background =element_rect(fill="grey95"),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 8),
        strip.text.y = element_text(colour = 'grey15'))

ggsave("FigS2.pdf", width = 8.5, height =9, units = "in")
```

#### Figure S3
```{r, fig.dim = c(8.5, 9)}
#HNI
#y-axis ordered by mean growth in control 

ggplot(nutODLim4combo, aes(Letter, reorder(CarbonSource, mean), fill=nutCat))+
  geom_tile()+
  facet_wrap(~SICA, nrow = 1, scales = "free_x")+
  scale_fill_manual(values = c("#EA725D","#F5BB62", "#85D1D1", "#5499C7", "grey99"),
                    labels = c("Suppressed Growth","Reduced Growth", "Increased Growth", "Induced Growth", "No Change"),
                    name = "")+
  labs(x="",
       y="")+
  theme_light()+
  theme(legend.position = "none",
        legend.text = element_text (size=8),
        axis.title.x = element_text (size = 8), 
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(), 
        strip.text.y.left = element_text(angle = 0),
        strip.background =element_rect(fill="grey95"),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 8),
        strip.text.y = element_text(colour = 'grey15'))

ggsave("FigS3.pdf", width = 8.5, height =9, units = "in")

```

#### Figure S4
```{r, fig.dim = c(8, 5)}

splitPlot <- split  %>%
    mutate(Treatment = case_when(Plot == "3" ~ "LNI",
                               Plot == "4" ~ "HNI"))

splitPlot$Treatment <- factor(splitPlot$Treatment, levels = c("LNI", "HNI"))

splitPlot %>%
  filter(comp != 'no change')%>%
ggplot(., aes(x=log, fill = comp))+
  geom_histogram(alpha=0.8, color =" white", size=0.2, breaks = seq(-6, 6, 0.5), aes(y=..count../sum(..count..)))+
  geom_vline(xintercept = 0, color = "white", size = 0.75)+
  geom_vline(xintercept = 0, lty = 3, color = "grey30")+
  theme_minimal()+
  facet_grid(Treatment~SICA)+
  labs(x = "",
       y = "proportion of interactions\n",
       fill = "")+
  scale_fill_manual(values = c("#76add2", "#ee8e7d"),
                    labels = c("decrease with SICA", "increase with SICA")
  )+
  theme(legend.position = "none")+
  scale_y_continuous(breaks = seq(0, 0.1, by = 0.01))

ggsave("FigS4.pdf", width = 8, height = 5, units = "in", device = cairo_pdf)
```
#### Figure S5
```{r, fig.dim=c(3.5, 5.5)}
#LNI Change in Growth per Carbon Source
heatChange3<- nutODLim3combo %>%
  mutate(testGroup = case_when(
    nutCat == "no_grow" ~ "W",
    nutCat != "no_grow" ~ "N")) %>%
  group_by(CarbonSource)%>%
  mutate(avgChange = round(mean(Change),3),
         group = "A",
         color = case_when(
           avgChange > 0 ~ "B",
           avgChange < 0 ~ "R",
           avgChange == 0 ~ "W"
         ))

#HNI Change in Growth per Carbon Source
heatChange4<- nutODLim4combo %>%
  mutate(testGroup = case_when(
    nutCat == "no_grow" ~ "W",
    nutCat != "no_grow" ~ "N")) %>%
  group_by(CarbonSource)%>%
  mutate(avgChange = round(mean(Change),3),
         group = "A",
         color = case_when(
           avgChange > 0 ~ "B",
           avgChange < 0 ~ "R",
           avgChange == 0 ~ "W"
         ))

#Combine
changeCombo  <- full_join(heatChange3, heatChange4, by = intersect(names(heatChange3), names(heatChange4)))

changeCombo %>%
  group_by(Plot, CarbonType)%>%
  summarise(changeAvg = mean(Change),
            sd = sd(Change),
            n=n(),
            se=sd/sqrt(n))%>%
  ggplot(., aes(changeAvg, reorder(CarbonType, -changeAvg), fill = Plot))+
  geom_col(position = position_dodge2(padding = 0.15),
           alpha=0.98,
           width = 0.8)+
      geom_errorbar(aes(xmin = changeAvg-se, 
                        xmax = -0.0009,
                        color = Plot),
                position = position_dodge(width = .85),
                size = 0.5, 
                width=0.2,
                show.legend = F)+
  geom_text(position = position_dodge(width = .85),
            aes(x = changeAvg -se, 
                label = round(changeAvg, 2), 
                hjust = 1.2,
                color = Plot),
            size = 3,
            fontface = "bold")+
  scale_fill_manual(values = c("#ee8e7d","#76add2"))+
  scale_color_manual(values = c("#ee8e7d","#76add2"))+
  theme_minimal()+
  scale_y_discrete(position = "right")+
  theme(legend.position = "none")+
  labs(x = "",
       y= "")+
  coord_cartesian(clip = "off",
                  xlim = c(-0.11, 0))


ggsave("FigS5.pdf", width = 3.5, height = 5.5, units = "in", device = cairo_pdf)
```

#### Figure S6
```{r, fig.dim = c(8.5, 9)}

#LNI Change in Growth per Carbon Source
ggplot(heatChange3, aes(Letter, reorder(CarbonSource, mean), fill=Change))+
  geom_tile()+
  facet_grid(CarbonType~SICA, scales = "free", space = "free")+
  scale_fill_gradientn(
    colours = colorRampPalette(c('#FD7446', '#FFFFFF', '#709AE1'))(100),
    breaks=c(-0.55,0, 0.55),
    limits=c(-0.55, 0.55),
    name=""
  )+
  geom_tile(data = ~ subset(., testGroup == "W"),
            fill="white")+
  coord_cartesian(clip = "off")+
  labs(x="",
       y="")+
  theme_light()+
  theme(legend.position = "none",
        axis.title.x = element_text (size = 8), 
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(), 
        strip.text.y.left = element_text(angle = 0),
        strip.background =element_rect(fill="grey95"),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 7),
        strip.text.y = element_blank(),
        strip.background.y=element_blank())

ggsave("FigS6.pdf", width = 8.5, height =9, units = "in")

#mean across carbon classes
heatChange3 %>%
  group_by(CarbonType)%>%
  summarise(mean = round(mean(avgChange),3))
```

#### Figure S7
```{r, fig.dim=c(8.5, 9)}

#HNI Change in Growth per Carbon Source
ggplot(heatChange4, aes(Letter, reorder(CarbonSource, mean), fill=Change))+
  geom_tile()+
  facet_grid(CarbonType~SICA, scales = "free", space = "free")+
  scale_fill_gradientn(
    colours = colorRampPalette(c('#FD7446', '#FFFFFF', '#709AE1'))(100),
    breaks=c(-0.55,0, 0.55),
    limits=c(-0.55, 0.55),
    name=""
  )+
  geom_tile(data = ~ subset(., testGroup == "W"),
            fill="white")+
  coord_cartesian(clip = "off")+
  labs(x="",
       y="")+
  theme_light()+
  theme(legend.position = "none",
        axis.title.x = element_text (size = 8), 
        axis.text.y = element_text(size = 6),
        axis.text.x = element_blank(), 
        strip.text.y.left = element_text(angle = 0),
        strip.background =element_rect(fill="grey95"),
        strip.text.x = element_text(colour = 'grey15', face = "bold", size = 7),
        strip.text.y = element_blank(),
        strip.background.y=element_blank())

ggsave("FigS7.pdf", width = 8.5, height =9, units = "in")

#mean across carbon classes
heatChange4 %>%
  group_by(CarbonType)%>%
  summarise(mean = round(mean(avgChange),3))

```

